---
title: "Homework 4"
author: "Kaylyn Zipp"
date: "3/1/2021"
output: html_document
---

1. Fit a Poisson regression model that assumes expected count is an interactive function of variables x1 and x2.
```{r}
Data<-read.csv("Homework 4 Data.csv")
head(Data)
fit<-glm(y ~ x1 * x2 , family = poisson, data = Data)
summary(fit)
```

2. Interpret the effect of variable x2 on the expected count when x2 is fixed at level 'b'. Verify your interpretation in R. 

$$
log(y_x) = \beta_0 + \beta_1 * X_1 +\beta_2*X_2b + \beta_3 *X_1 *X_2b
$$
$$
log(y_x) =\beta_0 + \beta_1*X_1 + \beta_2*1 + \beta_3*X_1*1
$$
$$
log(y_x) = (\beta_0 +\beta_2) + (\beta_1 + \beta_3) * X_1
$$

```{r}
betas<-coef(fit)

# log expected count at a 2
c_2<-(betas[1] + betas[3]) + (betas[2] + betas [4]) * 2
print(c_2)

# log expected count at 3
c_3<-(betas[1] + betas[3]) + (betas[2] + betas [4]) * 3
print(c_3)

effectvariablex1<-log(exp(c_3)/exp(c_2))
print(effectvariablex1)

confirmeffectofx1<-betas[4]+betas[2]
print(confirmeffectofx1)

```

3. Interpret the effect of variable x2 on the expected count when x1 is fixed at 1. Verify your interpretation in R. 

$$
log(y_x) = \beta_0 + \beta_1 * X_1 +\beta_2*X_2b + \beta_3 *X_1 *X_2b
$$
$$
log(y_x) = \beta_0 + \beta_1 * 1 +\beta_2*X_2b + \beta_3 *1 *X_2b
$$
$$
log(y_x) = (\beta_0 + \beta_1 ) + (\beta_2 + \beta_3 ) *X_2b
$$

When it is A 

$$
log(y_x) = (\beta_0+\beta_1)
$$
```{r}
c_a<-(betas[1] + betas[2])
print(c_a)

#When it is b
c_b<- (betas[1] + betas[2] ) + (betas[3] + betas[4] ) * 1
print(c_b)

effectvariablex2<-log(exp(c_b)/exp(c_a))
print(effectvariablex2)

confirmeffectofx1<-betas[3] + betas[4]
print(confirmeffectofx1)
```


4. Predict the expected count +/- 95% confidence intervals, over the observed range of values of x1, assuming x2 is fixed at level "b". 

```{r}
newdata <- data.frame(
  x1 = seq(min(Data$x1), max(Data$x1), length.out = 100),
  x2 = factor(x = rep('b', times = 100), 
              levels = c('a','b'))
)
prd <-predict.glm(object = fit, newdata = newdata, type = 'link', se.fit = T)
low <- exp(prd$fit - qnorm (0.975) *prd$se.fit)
high <- exp(prd$fit + qnorm (0.975) *prd$se.fit)

plot(y=exp(prd$fit), x = newdata$x1, xlab = 'x1', 
     ylab = 'Expected count', cex.axis = 1.5, cex.lab = 1.5,
     ylim = c(min(low), max(high)), type = 'l')
lines(x = newdata$x1, y = low, lty = 2)
lines(x = newdata$x1, y = high, lty = 2)
```


5. Predict the expected count +/- 95% confidence intervals, of levels "a" and "b", assuming x1 is fixed at its mean. 

```{r}

X1bar<-mean(Data$x1)
betas<-coef(fit)

?rep_len()
?factor

newdata <- data.frame(
  x1 = rep_len(x = X1bar, length.out = 100),
  x2 = sample(x = c("a","b"), replace =TRUE), 
              levels = c('a','b'))

prd <-predict.glm(object = fit, newdata = newdata, type = 'link', se.fit = T)
low <- exp(prd$fit - qnorm (0.975) * prd$se.fit)
high <- exp(prd$fit + qnorm (0.975) * prd$se.fit)

plot(y=exp(prd$fit), x = newdata$x2, xlab = 'x2', 
     ylab = 'Expected count', cex.axis = 1.5, cex.lab = 1.5,
     ylim = c(min(low), max(high)), type = 'l')

points(x = newdata$x2, y = low, lty = 2)
points(x = newdata$x2, y = high, lty = 2)



```




