---
title: "Homework 5"
author: "Kaylyn Zipp"
date: "4/1/2021"
output: html_document
---
1. Load data and place into an unmarkedFrameOccu object
```{r}
y<-read.csv('bobcat.csv')
PCovar<-read.csv("p covariates.csv")
PsiCovar<-read.csv('psi covariates.csv')

library(unmarked)
# converting detection / non-detection data into a matrix
sosp_mat <- as.matrix(y)
class(sosp_mat)
head(sosp_mat)
occu_data <- unmarkedFrameOccu(y = sosp_mat)


head(PCovar)

det_covs <- list(
people = data.frame(PCovar[, c('People1', 'People2','People3','People4', 'People5', 'People6', 'People7', 'People8', 'People9', 'People10', 'People11', 'People12', 'People13', 'People14', 'People15', 'People16', 'People17', 'People18', 'People19', 'People20', 'People21', 'People22', 'People23', 'People24', 'People25', 'People26', 'People27', 'People28', 'People29', 'People30', 'People31', 'People32', 'People33','People34','People35','People36','People37', 'People38', 'People39', 'People40', 'People41', 'People42', 'People43', 'People44', 'People45', 'People46', 'People47', 'People48', 'People49', 'People50', 'People51', 'People52','People53', 'People54', 'People55','People56', 'People57', 'People58', 'People59', 'People60', 'People61', 'People62', 'People63', 'People64', 'People65', 'People66', 'People67', 'People68', 'People69', 'People70', 'People71')]))


occu_data <- unmarkedFrameOccu(y = sosp_mat, # detection / non-detection
obsCovs = det_covs) # detection covariates

#Now I will add site level covariates

head(PsiCovar)
class(PsiCovar)

occu_data <- unmarkedFrameOccu(y = sosp_mat, # detection / non-detection
siteCovs = PsiCovar, # site-level covs
obsCovs = det_covs) # detection covariates
```

2. Fit an occupancy model that assumes conditional detection probability is a function of the number of people photographed at a site, and occupancy probability is a function of disturbance within 5km.

```{r}
fit <- occu(formula = ~ people ~ Dist_5km,

data = occu_data)

summary(fit)
```

3. Interpret the effect of the number of people photographed at each site on conditional detection probability of bobat. Verity your interpretation in R.


```{r}
coef(fit)
```
$$
y = \beta_0 + \beta_1*x1
$$

```{r}
#y= -3.01330391 + -0.06088908 *1
A <-(-3.01330391) + -0.06088908 *1

#y= -3.01330391 + -0.06088908 *2
B <- (-3.01330391) + -0.06088908 *2

B-A

plogis(B-A)

plogis(-0.06088908)
```

As the number of people seen at a location increases the log odds of seeing a bobcat decreases. The conditional detection probability decreases by 48.5%


4. Re-create the Wald test for the effect of disturbance within 5km. What is the null hypothesis? Do you reject or fail to reject this null hypothesis?

The null hypothesis would be that there is no effect of disturbance in 5 km on detection of bobcat at each camera. We reject the null hypothesis. 

```{r}
#Recreated Wald Test
summary(fit)
coef(fit)
# coefficient divided by standard error
w <- -23.643839	 / 4.7728768
w

2 * pnorm(-1 * abs(w))
```


5. Predict and plot the effect of disturbance within 5km on bobcat occupancy probability. Do this over the range of observed disturbance within 5km.


```{r}
new_psi <- data.frame(Dist_5km = seq(from = min(PsiCovar$Dist_5km),

to = max(PsiCovar$Dist_5km), length.out = 100))
predict <- predict(object = fit, newdata = new_psi,

type = 'state')

plot(y=predict$Predicted, x = new_psi$Dist_5km, xlab = 'Dist_5km', 
     ylab = 'Predicted', cex.axis = 1.5, cex.lab = 1.5, type = 'l')
lines(x= new_psi$Dist_5km, y = predict$lower, lty = 2)
lines(x = new_psi$Dist_5km, y = predict$upper,  lty = 2)
```


