---
title: "Homework 1"
author: "Kaylyn Zipp"
date: "February 7, 2021"
output:
  html_document: default
  html_notebook: default
---

1.Import the “beach.csv” dataset into R and examine the first few rows of data (5 points)

```{r}
beaches <- read.csv(file = 'beach.csv')
head(beaches)

```
To further understand the data I also 
```{r}
plot(x=beaches$Year, y=beaches$OpenBeach)
plot(x=beaches$BeachID, y=beaches$OpenBeach)
```

2.Fit a model that includes interactions between Year and BeachID. Report estimates of all coefficients
(5 points)

The annual change in open beach area created by BeachID

```{r}
fit <- lm(OpenBeach ~ Year * BeachID, data = beaches)
print(fit)
```

Equation
y = β0 + β1 × years + β2 × B + β3 × BeachID + β4 × years × BeachID + + β5 × years × BeachID

3. Predict the beach area at all three beaches in 1992 (5 points)

Area at Beach A

Beach A Dummy Coding = 0
Equation Breaks down to: y = β0 + β1 × years

36945.73 + -18.36 * 1992

```{r}
beachareaof_a_1992<-36945.73 + -18.36 * 1992
print(beachareaof_a_1992)
```

Area at Beach B

Beach B Dummy Coding = 1 for BeachIDB but 0 for BeachIDC
y = β0 + β1 × years + β2 × BeachIDB + β3 × BeachIDC + β4 × years × BeachIDB 
+ β5 × years × BeachIDC

Simplified

y = β0 + β1 × years + β2 × 1 +  β3 × 0 + β4 × years × 1 + β5 × years × 0

y = β0 + β1 × years + β2 × 1 + β4 × years × 1 

36945.73 + -18.36 * years + -34211.98 + 17.04 * years

36945.73 + -18.36 * 1992 + -34211.98 + 17.04 * 1992

```{r}
beachareaof_b_1992<-36945.73 + -18.36 * 1992 + -34211.98 + 17.04 * 1992
print(beachareaof_b_1992)
```

Area at Beach C

Beach C Dummy Coding = 0 for BeachIDB but 1 for BeachIDC
y = β0 + β1 × years + β2 × BeachIDB + β3 × BeachIDC + β4 × years × BeachIDB 
+ β5 × years × BeachIDC

Simplified
y = β0 + β1 × years + β3 × BeachIDC + β5 × years × BeachIDC

y = β0 + β1 × years + β3 × 1 + β5 × years × 1

36945.73 + -18.36 * years + -28776.11 + 14.37 * years

36945.73 + -18.36 * 1992 + -28776.11 + 14.37 * 1992

```{r}
beachareaof_c_1992 <- 36945.73 + -18.36 * 1992 + -28776.11 + 14.37 * 1992
print(beachareaof_c_1992)
```

4. Plot the predicted area of each beach across all years


```{r}
timespan<-seq(from = 1992, to = 2014, by = 1)
predictedbeachareaof_a<-36945.73 + -18.36 * timespan
predictedbeachareaof_b<-36945.73 + -18.36 * timespan + -34211.98 + 17.04 * timespan
predictedbeachareaof_c<-36945.73 + -18.36 * timespan + -28776.11 + 14.37 * timespan

plot(x = timespan, y = predictedbeachareaof_a, type = 'l', xlab = 'Years', ylab = 'Open Beach Area')  
lines(x = timespan, y=predictedbeachareaof_b, col = 'blue') 
lines(x = timespan, y = predictedbeachareaof_c, col = 'red')
legend('bottomleft', legend = c('Beach A', 'Beach B', 'Beach C'), col = c('red','blue','black'),pch = c('...'))
```



5. Interpret the rate of beach area change at each of the 3 beaches (5 points)

```{r}
BeachA <- lm(predictedbeachareaof_a ~ timespan)
print(BeachA)

BeachB <- lm(predictedbeachareaof_b ~ timespan)
print(BeachB)

BeachC <- lm(predictedbeachareaof_c ~ timespan)
print(BeachC)
```


All three of the beaches, A, B & C, are declining in open area over time steadily. Beach A is decreasing at the fastest rate, -18.36 units per year. Beach c is decreasing by a rate of 3.99 units per year. Beach b is decreasing the slowest, at a rate of -1.32 units per year. Based on the information provided if management is interested in stemming the lost of potential nesting sites then beach A should be focused on.   
