---
title: "Homework 7"
author: "Kaylyn Zipp"
date: "4/15/2021"
output: html_document
---

1. Load data and place into an unmarkedFrameOccu object

```{r setup, include=FALSE}
y <- read.csv("Bobcat.csv")
sosp_mat <- as.matrix(y)

p_covs <- read.csv('p covariates.csv')

det_covs <- list(
people = data.frame(p_covs[, c('People1', 'People2','People3','People4', 'People5', 'People6', 'People7', 'People8', 'People9', 'People10', 'People11', 'People12', 'People13', 'People14', 'People15', 'People16', 'People17', 'People18', 'People19', 'People20', 'People21', 'People22', 'People23', 'People24', 'People25', 'People26', 'People27', 'People28', 'People29', 'People30', 'People31', 'People32', 'People33','People34','People35','People36','People37', 'People38', 'People39', 'People40', 'People41', 'People42', 'People43', 'People44', 'People45', 'People46', 'People47', 'People48', 'People49', 'People50', 'People51', 'People52','People53', 'People54', 'People55','People56', 'People57', 'People58', 'People59', 'People60', 'People61', 'People62', 'People63', 'People64', 'People65', 'People66', 'People67', 'People68', 'People69', 'People70', 'People71')]))

site_covs<- read.csv('psi covariates.csv')

library(unmarked)
nmix_data<-unmarkedFramePCount(y = sosp_mat, 
siteCovs = site_covs, 
obsCovs = det_covs) 

```


2. Fit the following candidate set of models:
```{r}

library(AICcmodavg)

# Intercept ; Intercept
fit1 <- pcount(formula = ~ 1 ~ 1 ,nmix_data, K = 100,mixture = "P")
fit1_AIC<-AICc(mod = fit1, second.ord = F)
print(fit1_AIC)
# People ; Intercept
fit2 <- pcount(formula = ~ people ~ 1 ,nmix_data, K = 100,mixture = "P")
fit2_AIC<-AICc(mod = fit2, second.ord = F)
print(fit2_AIC)
# Intercept ; disturbance
fit3 <- pcount(formula = ~ 1 ~ Dist_5km ,nmix_data, K = 100,mixture = "P")
fit3_AIC<-AICc(mod = fit3, second.ord = F)
print(fit3_AIC)
summary(fit3)
#People ; Disturbance
fit4 <- pcount(formula = ~ people ~ Dist_5km,nmix_data, K = 100,mixture = "P")
fit4_AIC<-AICc(mod = fit4, second.ord = F)
print(fit4_AIC)
summary(fit4)

cand.set1<-list(P1 = fit1, P2 = fit2, P3 = fit3, P4 = fit4)
mods <- aictab(cand.set = cand.set1, second.ord = F)
head(mods)
```

3. Perform model selection with AIC. What is your top model? How do you know? Is there model selection
uncertainty?

The top model would be the intercept, disturbance model since it has the lowest AIC score. This would suggest that site specific variables are more influential than point. There is little difference between the intercept, disturbance and people, disturbance model creating some uncertainity. 

4. Average both the effect of people on detection, and disturbance on occupancy, over all models. Report
model-averaged slope coefficients and 95% confidence intervals.

```{r}
avg_type_1 <- modavgShrink(cand.set = cand.set1,
  parm = 'people',
  second.ord = F,
parm.type = "detect")
print(avg_type_1)

avg_type_2 <- modavgShrink(cand.set = cand.set1,
  parm = 'Dist_5km',
  second.ord = F,
parm.type = 'lambda')
print(avg_type_2)
```

Modeled Averaged Coefficients; 
People: -0.2 with 95% confidence intervals of -0.41, 0.37
Dist_5km: -23.71 with 95% confidence intervals of -32.4, -15.03


5. Obtain and plot model-averaged predictions of occupancy probability and detection probability. Average over all models, and make predictions over the observed range of each variable.

```{r}
library(unmarked)
library(AICcmodavg)

MinDist<-min(site_covs)
MaxDist<-max(site_covs)

new_dat <- data.frame(
Dist_5km = seq(from = MinDist, to = MaxDist, length.out = 100))

# calculating model averaged predictions
avg_prd <- modavgPred(cand.set = cand.set1,
newdata = new_dat,
second.ord = F,
parm.type = 'lambda')

avg_prd

plot(y=avg_prd$mod.avg.pred, x = new_dat$Dist_5km, xlab = 'Abundance Covariate', 
     ylab = 'Predicted', type = 'l')

lines(x= new_dat$Dist_5km, y = avg_prd$lower.CL, lty = 2)
lines(x= new_dat$Dist_5km, y = avg_prd$upper.CL, lty = 2)

p_covs[is.na(p_covs)] <- 0


# To find the Max and Min of the detection varaible

Max1<-max(p_covs$People1); Max2<-max(p_covs$People2); Max3<-max(p_covs$People3); Max4<-max(p_covs$People4); Max5<-max(p_covs$People5); Max6<-max(p_covs$People6);
Max7<-max(p_covs$People7); Max8<-max(p_covs$People8); Max9<-max(p_covs$People9); 
Max10<-max(p_covs$People10)

Max11<-max(p_covs$People11); Max12<-max(p_covs$People12); Max13<-max(p_covs$People13)
Max14<-max(p_covs$People14); Max15<-max(p_covs$People15); Max16<-max(p_covs$People16)
Max17<-max(p_covs$People17); Max18<-max(p_covs$People18); Max19<-max(p_covs$People19)

Max20<-max(p_covs$People20); Max21<-max(p_covs$People21); Max22<-max(p_covs$People22)
Max23<-max(p_covs$People23); Max24<-max(p_covs$People24); Max25<-max(p_covs$People25)
Max26<-max(p_covs$People26); Max27<-max(p_covs$People27); Max26<-max(p_covs$People26)
Max27<-max(p_covs$People27); Max28<-max(p_covs$People28); Max29<-max(p_covs$People29)

Max30<-max(p_covs$People30); Max31<-max(p_covs$People31); Max32<-max(p_covs$People32)
Max33<-max(p_covs$People33); Max34<-max(p_covs$People34); Max35<-max(p_covs$People35)
Max36<-max(p_covs$People36); Max37<-max(p_covs$People37); Max36<-max(p_covs$People36)
Max37<-max(p_covs$People37); Max38<-max(p_covs$People38); Max39<-max(p_covs$People39)

Max40<-max(p_covs$People40); Max41<-max(p_covs$People41); Max42<-max(p_covs$People42)
Max43<-max(p_covs$People43); Max44<-max(p_covs$People44); Max45<-max(p_covs$People45)
Max46<-max(p_covs$People46); Max47<-max(p_covs$People47); Max46<-max(p_covs$People46)
Max47<-max(p_covs$People47); Max48<-max(p_covs$People48); Max49<-max(p_covs$People49)

Max50<-max(p_covs$People50); Max51<-max(p_covs$People51); Max52<-max(p_covs$People52)
Max53<-max(p_covs$People53); Max54<-max(p_covs$People54); Max55<-max(p_covs$People55)
Max56<-max(p_covs$People56); Max57<-max(p_covs$People57); Max56<-max(p_covs$People56)
Max57<-max(p_covs$People57); Max58<-max(p_covs$People58); Max59<-max(p_covs$People59)

Max60<-max(p_covs$People60); Max61<-max(p_covs$People61); Max62<-max(p_covs$People62)
Max63<-max(p_covs$People63); Max64<-max(p_covs$People64); Max65<-max(p_covs$People65)
Max66<-max(p_covs$People66); Max67<-max(p_covs$People67); Max66<-max(p_covs$People66)
Max67<-max(p_covs$People67); Max68<-max(p_covs$People68); Max69<-max(p_covs$People69)

Max70<-max(p_covs$People70); Max71<-max(p_covs$People71)

new_dat2 <- data.frame(
People1 = seq(from = 0, to = Max1, length.out = 100), People2 = seq(from = 0, to = Max2, length.out = 100), People3 = seq(from = 0, to = Max3, length.out = 100),People4 = seq(from = 0, to = Max4, length.out = 100), People5 = seq(from = 0, to = Max5, length.out = 100), People6 = seq(from = 0, to = Max6, length.out = 100), People7 = seq(from = 0, to = Max7, length.out = 100), People8 = seq(from = 0, to = Max8, length.out = 100),People9 = seq(from = 0, to = Max9, length.out = 100), People10 = seq(from = 0, to = Max10, length.out = 100), People11 = seq(from = 0, to = Max11, length.out = 100), People12 = seq(from = 0, to = Max12, length.out = 100), People13 = seq(from = 0, to = Max13, length.out = 100), People14 = seq(from = 0, to = Max14, length.out = 100),People15 = seq(from = 0, to = Max15, length.out = 100),People16 = seq(from = 0, to = Max16, length.out = 100),People17 = seq(from = 0, to = Max17, length.out = 100),People18 = seq(from = 0, to = Max18, length.out = 100),People19 = seq(from = 0, to = Max19, length.out = 100),People20 = seq(from = 0, to = Max20, length.out = 100),People21 = seq(from = 0, to = Max21, length.out = 100),People22 = seq(from = 0, to = Max22, length.out = 100),People23 = seq(from = 0, to = Max23, length.out = 100),People24 = seq(from = 0, to = Max24, length.out = 100),People25 = seq(from = 0, to = Max25, length.out = 100),People26 = seq(from = 0, to = Max26, length.out = 100),People27= seq(from = 0, to = Max27, length.out = 100),People28= seq(from = 0, to = Max28, length.out = 100),People29= seq(from = 0, to = Max29, length.out = 100),People30 = seq(from = 0, to = Max30, length.out = 100),People31 = seq(from = 0, to = Max31, length.out = 100),People32= seq(from = 0, to = Max32, length.out = 100),People33 = seq(from = 0, to = Max33, length.out = 100),People34 = seq(from = 0, to = Max34, length.out = 100),People35 = seq(from = 0, to = Max35, length.out = 100),People36 = seq(from = 0, to = Max36, length.out = 100),People37 = seq(from = 0, to = Max37, length.out = 100),People38 = seq(from = 0, to = Max38, length.out = 100),People39 = seq(from = 0, to = Max39, length.out = 100),People40 = seq(from = 0, to = Max40, length.out = 100),People41 = seq(from = 0, to = Max41, length.out = 100),People42 = seq(from = 0, to = Max42, length.out = 100),People43 = seq(from = 0, to = Max43, length.out = 100),People44 = seq(from = 0, to = Max44, length.out = 100),People45 = seq(from = 0, to = Max45, length.out = 100),People46 = seq(from = 0, to = Max46, length.out = 100),People47 = seq(from = 0, to = Max47, length.out = 100),People48 = seq(from = 0, to = Max48, length.out = 100),People49 = seq(from = 0, to = Max49, length.out = 100),People50 = seq(from = 0, to = Max50, length.out = 100),People51 = seq(from = 0, to = Max51, length.out = 100),People52 = seq(from = 0, to = Max52, length.out = 100),People53 = seq(from = 0, to = Max53, length.out = 100),People54 = seq(from = 0, to = Max54, length.out = 100),People55 = seq(from = 0, to = Max55, length.out = 100),People56 = seq(from = 0, to = Max56, length.out = 100),People57 = seq(from = 0, to = Max57, length.out = 100),People58 = seq(from = 0, to = Max58, length.out = 100),People59 = seq(from = 0, to = Max59, length.out = 100),People60 = seq(from = 0, to = Max60, length.out = 100),People61 = seq(from = 0, to = Max61, length.out = 100),People62 = seq(from = 0, to = Max62, length.out = 100),People63 = seq(from = 0, to = Max63, length.out = 100),People64 = seq(from = 0, to = Max64, length.out = 100),People65 = seq(from = 0, to = Max65, length.out = 100),People66 = seq(from = 0, to = Max66, length.out = 100),People67 = seq(from = 0, to = Max67, length.out = 100),People68 = seq(from = 0, to = Max68, length.out = 100),People69 = seq(from = 0, to = Max69, length.out = 100),People70 = seq(from = 0, to = Max70, length.out = 100), People71 = seq(from = 0, to = Max71, length.out = 100))

#new data

new_dat3 <- data.frame(
people = seq(from = 0, to = 5.6, length.out = 100)) 

                                     
avg_prd2<- modavgPred(cand.set = cand.set1,
newdata = new_dat3,
second.ord = F,
parm.type = "detect")

plot(y=avg_prd2$mod.avg.pred, x = new_dat3$people, xlab = 'Abundance Covariate', 
     ylab = 'Predicted', type = 'l', ylim = c(0,.3))

lines(x= new_dat3$people, y = avg_prd2$lower.CL, lty = 2)
lines(x= new_dat3$people, y = avg_prd2$upper.CL, lty = 2)



```






