---
title: "Homework6"
author: "Kaylyn Zipp"
date: "4/11/2021"
output: html_document
---

1. Load data and place into an unmarkedFramePCount object

```{r}
y<-read.csv('count.csv')
head(y)

p_covs <- read.csv('obs_covs.csv')
head(p_covs)

det_covs <- list(
j = data.frame(p_covs[, c('j1', 'j2','j3')]))

site_covs <- read.csv('site_covs.csv')
head(site_covs)
site_covs$x2<-factor(site_covs$x2)

library(unmarked)
sosp_mat <- as.matrix(y)
nmix_data <- unmarkedFramePCount(y = sosp_mat, siteCovs = site_covs, obsCovs = det_covs)
```

2. Fit an N-mixture model that assumes conditional detection probability is a function of the detection
covariate provided, and expected abundance is a additive function of variables x1 and x2.
```{r}
fit <- pcount(formula = ~ j ~ x1 + x2, data = nmix_data, K = 100)
summary(fit)
```

3. Interpret the effect of x1 on the expected count at each site. Verity your interpretation in R.
```{r}
b<-coef(fit)

exp_1<-b[1]+b[2]*1
exp_2<-b[1]+b[2]*2
diff<-exp_1 - exp_2
print(diff)

b[2]

exp(diff)
exp(b[2])

```

4. Predict and plot the effect of the supplied detection covariate. Do this over the range of this covariate.

```{r}
Min<-min(p_covs)
Max<-max(p_covs)
new_p <- data.frame( j = seq(from = Min, to = Max, length.out = 100))

prd_p <- predict(object = fit, newdata = new_p, type = 'det')

plot(y=prd_p$Predicted, x = new_p$j, xlab = 'Detection Covariate', 
     ylab = 'Predicted', type = 'l')

lines(x= new_p$j, y = prd_p$lower, lty = 2)
lines(x= new_p$j, y = prd_p$upper, lty = 2)
```


5. Use contrasts to compare expected abundance between all pairwise levels of variable x2. Obtain p-values associated with each contrast and tell me whether you reject or fail to reject each null hypothesis tested.

```{r}
x <- matrix(
c(0, 0, 1, -1, 0,
0, 0, 1, 0, -1,
0, 0, 0, 1, -1),
nrow = 3, byrow = T
)
x
lin_com <- linearComb(obj = fit, coefficients = x, type = 'state')
lin_com

w <- coef(lin_com) / SE(lin_com)
w
pvalues<-2 * pnorm(-1 * abs(w))
```

We fail to reject the null hypothesis when comparing level b to level c
```{r}
pvalues[1]
```

We reject the null hypothesis when comparing level b to level d
```{r}
pvalues[2]
```

We reject the null hypothesis when comparing level c to level d
```{r}
pvalues[3]
```

